---
title: "06-project_geomorph"
author: "Andrew G. Brown"
date: "2020-05-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Defining spatial extent of "low foothills"

 - _Low relief_: vertical distance from channel (SAGA GIS; 10m resolution) less than 30m
 
 - _Low gradient_: slope gradient (SAGA GIS; 10m resolution) less than 30%
 
 - _Low precipitation_: 
 
   - 18" is low cutoff for F200 (18 - 25") and midrange for R163 (14 - 22")
   
   - 22" is low cutoff for F201 (22 - 31") and high cutoff for R163 (14 - 22")
   
   - 25" is high cutoff for F200 (18 - 25")
 
 - _Low tree cover_: low blue oak probability / spectral derivative? Most are NLCD "grassland" areas (TODO)

```{r}
# SETUP
slope.threshold        <- 30            # % slope
vdistchannel.threshold <- 30            # meters
precip.thresholds      <- c(18, 22, 25) # inches of rain
```

```{r, echo=FALSE, warning=FALSE}
library(raster)
vdistchannel <- raster('E:/CA649/Geodata/derived/vdistchannel.tif')
slope <- tan(raster('E:/CA649/Geodata/derived/slope.tif'))*100

lowrelief <- (vdistchannel <= vdistchannel.threshold)
lowgradient <- (slope <= slope.threshold)
lowhills <- (lowrelief + lowgradient)

plot(lowhills, main="Low Relief + Low Gradient in Project Extent\n(0 = neither; 1 = either; 2 = both)")

writeRaster(lowhills == 2, filename="E:/CA649/Geodata/derived/lowhills.tif", overwrite=T)

precip <- raster("E:/CA649/Geodata/derived/precip_mm.tif")

bound <- rgdal::readOGR("E:/CA649/Geodata/source/CA649_b.shp")
bound <- spTransform(bound, CRS(proj4string(lowhills)))

for(i in 1:length(precip.thresholds)) {
  lowprecip <- precip < (precip.thresholds[i] * 25.4)
  
  res <- lowprecip + lowhills == 3
  plot(res, main=paste0("Precip < ",precip.thresholds[i],
                        "\"; Slope < ",slope.threshold,
                        "%; Vertical dist. from channel < ",
                        vdistchannel.threshold,"m"))
  
  writeRaster(res, 
              filename=paste0("E:/CA649/Geodata/derived/lowhills_", precip.thresholds[i], ".tif"),
              overwrite=TRUE)
  lines(bound)
}
```

